### 2025-04-15 reports/paper/paper.qmd
some revisions

---
# fontsize: 11pt
# mainfont: arial
format:
 # html: default
 docx:
    number-sections: false
    global-numbering: true
    reference-doc: "../../../../HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
notes-after-punctuation: true
execute:
  freeze: auto
editor_options: 
  chunk_output_type: inline
editor: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=F, results = 'asis')
library(bibtex)
biblio <- bibtex::read.bib('~/Box Sync/papers/EnvEpi.bib')
library(flextable)
library(data.table)
library(ggplot2); library(magrittr)
```

# Abstract

(Limit: 200 words)

@@

\newpage

# Introduction

@@

# Methods

@@

::: {#fig-rules layout-ncol="2"}
![](resources/images/intervention-density-witnin-combos-a.png){#fig-1a
width="3.5in"}
![](resources/images/intervention-density-witnin-combos-b.png){#fig-1b
width="3.5in"}
![](resources/images/intervention-density-witnin-combos-c.png){#fig-1c
width="3.5in"}

## Study population

@@

## Outcome and potential confounders

@@

## Exposure

@@

## Hazard-extended ICE parametric g-formula

@@

We estimated cumulative incidence under no intervention on soluble MWF
and under the supportable interventions based on the five selected
target exposure limits. We contrasted the cumulative incidence under the
supportable interventions to that under no intervention on exposure by
computing the number of cases averted and the cumulative incidence
ratios. Confidence intervals were computed using the basic nonparametric
bootstrap with 1000 Monte Carlo samples from the population at the start
of follow-up. All the necessary scripts used to reproduce the analyses
are available on [GitHub](https://github.com/kvntchn/gm-nhl-ice).

@@

# Results

@@

```{r}
tab1.sub |> 
  data.frame() |> flextable() |>
  delete_part(part = "header") |>
  add_header_row(values = c("", "Study population", "", "NHL cases"),
                 colwidths = c(1, 2, 1, 2)) |>
  hline_top(part = "header") |>
  hline_top() |> hline_bottom() |>
  hline(which((grepl("[0-9]\\)", tab1.sub$spread))[-1])[1]) |> width(4, 0.1) |>
  width(1, 2.2) |> width(c(3, 6), 1.1) |> width(c(2, 5), 0.65) |>
  flextable::compose(
    i = grep("Cumulative exposure", rownames(tab1.sub)), j = 1,
    value = as_paragraph("Cumulative exposure to MWF (mg/m",
                         as_sup("3"), "-years)")) |>
  flextable::align(j = 2:6, align = "center", part = "all") |>
  footnote(i = grep("Plant$", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Plant of longest employment duration among those who worked at multiple plants"),
    ref_symbols = "a", inline = T) |>
  footnote(i = grep("Ever exposed|time off", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Lagged 20 years"),
    ref_symbols = "b", inline = T) |>
  # footnote(i = grep("work", rownames(tab1.sub)), j = 1, value = as_paragraph(
  #   " Among those who left work by December 31, 1994"),
  #   ref_symbols = "c", inline = T) |>
  # footnote(i = grep("death", rownames(tab1.sub)), j = 1, value = as_paragraph(
  #   " Among deceased"),
  #   ref_symbols = "d", inline = T) |>
  footnote(i = grep("Cumulative exposure", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Among ever-exposed individuals, lagged 20 years."),
    ref_symbols = "d", inline = T) |>
  add_footer_row(values = "NHL: non-Hodgkin lymphoma.", colwidths = 6) |>
  padding(padding = 1.5, part = "all") |>
  padding(
    i = which(tab1.sub_hspace),
    j = 1,
    padding.left = 20
  ) |>
  merge_at(i = grep("Cumulative exposure", rownames(tab1.sub))) |>
  hline_bottom(part = "footer") |> font(fontname = 'arial') |>
	set_caption(
		caption = paste0(
			"Summary of population characteristics. Statistics shown above ",
			"the horizontal line are count (%). Those shown below are median ",
			"(quartile 1, quartile 3).")
	)
```

@@

![Figure 2. Median annual average daily exposure to soluble, straight,
and synthetic metalworking fluids among exposed workers over
time.](../../resources/images/exposure.png){width="6.5in"}

@@

The observed number of NHL cases over the 30-year follow-up period was
339. Table 2 presents estimates of the counterfactual number of cases,
number of cases averted, and cumulative incidence ratios contrasting
supportable interventions on exposure to soluble MWF based on different
target exposure limits and no censoring. Under an intervention
eliminating censoring, the estimated number of cases was 502 (95% CI:
439, 555). Interventions based on stronger target exposure limits on
soluble MWF resulted in monotonically stronger reductions in the
estimated cumulative incidence of NHL. Setting the target exposure limit
at the NIOSH REL 0.5 mg/m^3^ would have averted 124 (95% CI: 66, 202)
NHL cases.

```{r, eval=F}
tikzDevice::tikz(here::here("reports/private/SER (2023)/resources/rd.tex"),
                 width = 3.75, height = 2.25,
								 documentDeclaration = "\\documentclass{beamer}",
                 bareBones = T)
contrasts.tab[,.(
     label,
     # `Person-years intervened (%)` = clean_percent_py(variable),
     `Cases averted`    = rd * N,
     rd_lower,
     rd_upper,
     result = paste0(
     	round(rd * N),
     	" (", round(rd_lower), ", ", round(rd_upper), ")")
   )][c(6, 3, 2, 5, 4, 1),] -> gg.tab 
gg.tab[label != "None"] |> ggplot(
    aes(x = factor(
    	label,
    	levels = c("0.05", "0.25", "0.5", "1.0", "2.0",  "None"),
    	labels = c("0.05", "0.25", "0.50", "1.00", "2.00",  "None")),
        y = `Cases averted`, ymin = rd_lower, ymax = rd_upper)
  ) +
	geom_label(aes(y = 270, label = result), size = 8/.pt, 
						 label.padding = unit(5, "pt")) +
  geom_errorbar(linewidth = 4 / .pt, width = 0.15, color = '#003c61') +
  geom_point(size = 4 / .pt, color = '#003c61') + 
  geom_hline(yintercept = 0, lty = 2, , alpha = 0.7) +
  labs(
    y = "Cases averted",
    x = "Target limit (mg/m$^3$)") +
  mytheme +
  coord_flip(ylim = c(-10, 320)) +
  theme(panel.grid = element_blank(),
  			axis.text = element_text(size = 9, colour = 'black'),
  			panel.border = element_rect(colour = "black", size = 0.75),
  			axis.title.y = element_text(size = 9, margin = margin(r = 6)),
  			axis.title.x = element_text(size = 9, margin = margin(t = 6)),
  			plot.margin = margin(l = 6, r = 6)
  )
dev.off()
```

```{r}
contrasts.tab[,.(
     label,
     # `Person-years intervened (%)` = clean_percent_py(variable),
     `Cases`  = clean_risk(variable, 0, N),
     `(95% CI)`      = clean_risk.ci(variable, 0, N),
     `Cases averted`    = round(rd * N),
     `(95% CI1)`        = paste0(
       "(",
       round(rd_lower), ", ",
       round(rd_upper), ")"),
     CIR              = sprintf("%.2f", round(rr, 2)),
     `(95% CI2)`   = clean_rr.ci(rr, rr_lower, rr_upper)
   )][c(6, 3, 2, 5, 4, 1),] -> sol.tab
sol.tab[label == "None", 4:7] <- NA

# save(sol.tab, mwf.flextable, file = here::here('resources', 'sol.tab.rdata'))

sol.tab |> flextable() |>
  flextable::compose(
    j = 1,
    value = as_paragraph("Target exposure limit (mg/m", as_sup("3"), ")"),
    part = "header"
  ) |>
  set_header_labels(
    `(95% CI1)` = "(95% CI)",
    `(95% CI2)` = "(95% CI)"
  ) |> mwf.flextable() |>
  width(width = c(1.1, 0.5, 0.9, 0.55, 0.9, 0.5, 0.9)) |>
  add_footer_row(values = paste(
    "CIR: cumulative incidence ratio. MWF: metalworking fluid",
    sep = "\n"), colwidths = 7) |>
  hline_bottom(part = "footer") |>
	set_caption(caption = paste0(
		"Estimates of the counterfactual number of cases, ",
		"number of cases averted ",
		"and cumulative incidence ratios ",
		"contrasting supportable ",
		"interventions on annual average daily exposure to soluble MWF ",
		"to no intervention on exposure."
	))

# Discussion

# Conclusions

# Citations

>>>>

---
# title: Hypothetical interventions on workplace exposure with guaranteed positivity
fontsize: 11pt
header-includes:
- \allowdisplaybreaks
- \usepackage{setspace}\setstretch{1.1}
- \usepackage{etoolbox}
  \newcommand{\zerodisplayskips}{
    \setlength{\abovedisplayskip}{2pt}
    \setlength{\belowdisplayskip}{2pt}
    \setlength{\abovedisplayshortskip}{1pt}
    \setlength{\belowdisplayshortskip}{1pt}}
  \appto{\normalsize}{\zerodisplayskips}
  \appto{\small}{\zerodisplayskips}
  \appto{\footnotesize}{\zerodisplayskips}
- \usepackage{pgf}\usepackage{pgfplots}\usepackage{tikz}
- \usetikzlibrary{graphs, arrows, automata, shadings}
- \setlength{\floatsep}{0pt}
- \usepackage{enumitem}\setlist[enumerate]{leftmargin=*}
- \usepackage{verbatim}
- \makeatletter
  \patchcmd{\@verbatim}{\verbatim@font}{\singlespacing\small\verbatim@font}{}{}
- \RedeclareSectionCommand[runin=false, beforeskip=-1sp,
  afterskip=-0.1\baselineskip]{section}
- \RedeclareSectionCommand[runin=false, beforeskip=-1sp,
  afterskip=-0.1\baselineskip]{subsection}
- \newcommand\blfootnote[1]{
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}
  \addtocounter{footnote}{-1}
  \endgroup}
- \setlength{\columnsep}{2em}
- \setlength{\tabcolsep}{2pt}
- \usepackage{dsfont}
- \usepackage[section]{placeins}
latex-output-dir: '..'
format:
  pdf:
    documentclass: scrartcl
    # classoption: twocolumn
    geometry: 
      - margin=1.1in
      - footskip=0.35in
    keep-tex: true
  # docx:
  #   number-sections: false
  #   global-numbering: true
  #   reference-doc: "../../../../HeadRs/rmd_templates/style-guide-serif.docx"
tbl-cap-location: top
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
notes-after-punctuation: true
csl: "`r path.expand('~/HeadRs/csl/de-gruyter.csl')`"
fig-cap-location: top
crossref:
  lof-title: "List of figures"
  lot-title: "List of tables"
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Figure S
      space-before-numbering: false
      latex-list-of-description: Supplemental Figure
      caption-location: top
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Table S
      space-before-numbering: false
      latex-list-of-description: Supplemental Table
      caption-location: top
execute:
  freeze: auto
editor_options: 
  chunk_output_type: inline
---

```{r, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=F, results = 'asis')
source("~/headrs/00-my-theme.R")
library(xtable)
library(flextable)
set_flextable_defaults(font.size = 10, font.family = "Helvetica Neue")
library(data.table)
library(ggplot2); library(magrittr)
```

# Hypothetical interventions on workplace exposure with guaranteed positivity

## Abstract

@@

{{< pagebreak >}}

## Introduction

@@

## Methods

@@

### Study population

@@

### Outcome and potential confounders

@@

### Exposure

@@

### Hazard-extended ICE parametric g-formula

@@

We estimated cumulative incidence under no intervention on soluble MWF
and under the supportable interventions based on the five selected
target exposure limits. We contrasted the cumulative incidence under the
supportable interventions to that under no intervention on exposure by
computing the risk difference per 10,000 and the cumulative incidence
ratios. Confidence intervals were computed using the basic nonparametric
bootstrap with 1000 Monte Carlo samples from the population at the start
of follow-up. All the necessary scripts used to reproduce the analyses
are available on [GitHub](https://github.com/kvntchn/gm-nhl-ice).

## Results

The observed number of NHL cases over the 30-year follow-up period was
339. Table 2 presents estimates of the counterfactual number of cases,
risk difference per 10,000, and cumulative incidence ratios contrasting
supportable interventions on exposure to soluble MWF based on different
target exposure limits and no censoring. Under an intervention
eliminating censoring, the estimated number of cases was 502 (95% CI:
439, 555). Interventions based on stronger target exposure limits on
soluble MWF resulted in monotonically stronger reductions in the
estimated cumulative incidence of NHL. Setting the target exposure limit
at the NIOSH REL 0.5 mg/m^3^ would have averted 124 (95% CI: 66, 202)
NHL cases.

## Discussion

## Conclusions

## References

{{< pagebreak >}}

## Tables and figures

:::{#fig-rules layout-ncol=2}
![](resources/intervention-density-witnin-combos-a.pdf){#fig-1a}

![](resources/intervention-density-witnin-combos-b.pdf){#fig-1b}

![](resources/intervention-density-witnin-combos-c.pdf){#fig-1c}

@@

@@

:::{#tbl-1}
```{r}
if (!knitr::is_latex_output()) {
tab1.sub |> 
  data.frame() |> flextable() |>
  delete_part(part = "header") |>
  add_header_row(values = c("", "Study population", "", "NHL cases"),
                 colwidths = c(1, 2, 1, 2)) |>
  hline_top(part = "header") |>
  hline_top() |> hline_bottom() |>
  hline(which((grepl("[0-9]\\)", tab1.sub$spread))[-1])[1]) |> width(4, 0.1) |>
  width(1, 2.2) |> width(c(3, 6), 1.1) |> width(c(2, 5), 0.65) |>
  flextable::compose(
    i = grep("Cumulative exposure", rownames(tab1.sub)), j = 1,
    value = as_paragraph("Cumulative exposure to MWF (mg/m",
                         as_sup("3"), "-years)")) |>
  flextable::align(j = 2:6, align = "center", part = "all") |>
  footnote(i = grep("Plant$", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Plant of longest employment duration among those who worked at multiple plants"),
    ref_symbols = "a", inline = T) |>
  footnote(i = grep("Ever exposed|time off", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Lagged 20 years"),
    ref_symbols = "b", inline = T) |>
  # footnote(i = grep("work", rownames(tab1.sub)), j = 1, value = as_paragraph(
  #   " Among those who left work by December 31, 1994"),
  #   ref_symbols = "c", inline = T) |>
  # footnote(i = grep("death", rownames(tab1.sub)), j = 1, value = as_paragraph(
  #   " Among deceased"),
  #   ref_symbols = "d", inline = T) |>
  footnote(i = grep("Cumulative exposure", rownames(tab1.sub)), j = 1, value = as_paragraph(
    " Among ever-exposed individuals, lagged 20 years."),
    ref_symbols = "d", inline = T) |>
  add_footer_row(values = "NHL: non-Hodgkin lymphoma.", colwidths = 6) |>
  padding(padding = 1.5, part = "all") |>
  padding(
    i = which(tab1.sub_hspace),
    j = 1,
    padding.left = 20
  ) |>
  merge_at(i = grep("Cumulative exposure", rownames(tab1.sub))) |>
  hline_bottom(part = "footer")
} else {
	rownames(tab1.sub) <- NULL
	tab1.sub <- tab1.sub[,-4]
	tab1.sub[tab1.sub_hspace, "variable"] <- paste0(
		"\\hspace{2em}", tab1.sub[tab1.sub_hspace, "variable"])
	tab1.sub$spread <- sanitize(tab1.sub$spread)
	tab1.sub$spread.1 <- sanitize(tab1.sub$spread.1)
	names(tab1.sub) <- c(
		"%",
		"Count", "(\\%)",
		"Count", "(\\%)")
	print(xtable(tab1.sub,
							 align = "clrlrl"),
				add.to.row = list(
					pos = list(-1, 1, 16),
					command = c(
						paste0(
							"\\toprule & \\multicolumn{2}{c}{Study population}", 
							" & \\multicolumn{2}{c}{NHL Cases} \\\\ \n",
							"\\cmidrule(lr){2-3}\\cmidrule(lr){4-5}\n"),
						"\\midrule & Count & (\\%)     & Count & (\\%) \\\\\n\\midrule\n",
						"\\midrule & Median & (Q1, Q3) & Median & (Q1, Q3) \\\\\n\\midrule\n"
					)
				),
				hline.after = nrow(tab1.sub)
	)
}

Summary of population characteristics. Statistics shown above
the horizontal line are count (%). Those shown below are median
(quartile 1, quartile 3).

:::

@@

:::{#fig-exposure}
![](../../resources/images/exposure.png){width="6.5in"}

Median annual average daily exposure to soluble, straight,
and synthetic metalworking fluids among exposed workers over
time.
:::

@@

```{r}
contrasts.tab[,.(
     label,
     # `Person-years intervened (%)` = clean_percent_py(variable),
     `Cases`  = clean_risk(variable, 0, N),
     `(95% CI)`      = clean_risk.ci(variable, 0, N),
     `RD`    = round(rd * 10000),
     `(95% CI1)`        = paste0(
       "(",
       round(rd_lower), ", ",
       round(rd_upper), ")"),
     CIR              = sprintf("%.2f", round(rr, 2)),
     `(95% CI2)`   = clean_rr.ci(rr, rr_lower, rr_upper)
   )][c(6, 3, 2, 5, 4, 1),] -> sol.tab
sol.tab[label == "None", 4:7] <- NA

# save(sol.tab, mwf.flextable, file = here::here('resources', 'sol.tab.rdata'))
```

::: {#tbl-rd}
```{r}
if (!knitr::is_latex_output()) {
sol.tab |> flextable() |>
  flextable::compose(
    j = 1,
    value = as_paragraph("Target exposure limit (mg/m", as_sup("3"), ")"),
    part = "header"
  ) |>
  set_header_labels(
    `(95% CI1)` = "(95% CI)",
    `(95% CI2)` = "(95% CI)"
  ) |> mwf.flextable() |>
  width(width = c(1.1, 0.5, 0.9, 0.55, 0.9, 0.5, 0.9)) |>
  add_footer_row(values = paste(
    "CIR: cumulative incidence ratio. MWF: metalworking fluid",
    sep = "\n"), colwidths = 7) |>
  hline_bottom(part = "footer")
} else {
	names(sol.tab) <- gsub("CI.*)", "CI)", names(sol.tab))
	names(sol.tab) <- sanitize(names(sol.tab))
	names(sol.tab)[1] <- "Target exposure limit (mg/m\\textsuperscript{3})"
	xtable(sol.tab, align = c("clrlrlrl"))
}
```

Estimates of the counterfactual number of cases, number of cases 
averted, and cumulative incidence ratios (CIR) contrasting supportable 
interventions on annual average daily exposure to soluble MWF
to no intervention on exposure.
:::
